---
title: "Project – Data Science Skills"
author: "Team name : CUNY VERSE"
date: "March 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Team Members
  Mia Siracusa ;
  Javern;
  Kleber Perez;
  Yohannes Getahun Deboch
  
##  Introduction

Data Science Skills Project focused on Exploratory Analysis of  US Chrnoic Disease Indicators DatA Set
  
##  Objective of the Project

The objective of this project is to practice soft skill working in a virtual taem. During the project we practised collaborating, knowledge sharing and problem solving in a team remotely .
Load the libraries
```{r}
library(tidyverse)
library(Amelia)

```
# Approach

For this data science skills project after having detiled discussion, we have decided to use US chronic disease indicators data proposed by Kleber (one of the team members)
Using this data set we followed the following approach:
1. Means of Communication: We created a Whatsapp group where we can share documents and make a group call to discuss how we should do the project. For the purpose of sharing data sets we have utilized Github and Rpubs . 
2. Separate the work among the team members.
3. Each individual do their part of the work.
4. Combine everyone's work to make a final work.

# Motivation

Our motivation behind working as a team is due to the necessity of being able to effectivelly collaborate in a virtual environment . Working in a virtual team is essentail  as it will enhance one of the basic data science soft skills( which being effective communiction and collaboration). Nowadays people prefer more flexible work schedule and working from home due to the inherent disire of maintaing work and life balance. This project has enabled us to practice how to mantain  work and life balance working in a virtual team.

# Data information
For this data Science skills project we're using us chronic disease indicators data set.

# Data source
The data was downloaded from the following website in csv format.
https://chronicdata.cdc.gov/Chronic-Disease-Indicators/U-S-Chronic-Disease-Indicators-CDI-/g4ie-h725

# Data on google drive
https://drive.google.com/file/d/14lQCOt5gHB6lk9995hsgB5cxgH8BUMU1/view?usp=sharing

# How to load it
We've loaded the data using read.csv and for indicating missing values empty string was used as an identifier.
Load the data
```{r}
url <- "https://doc-04-28-docs.googleusercontent.com/docs/securesc/4gka1akk08hvpslffei635a9a065i68g/li4rhokegsnad94d1cg1n01qjkvbqgpc/1553011200000/07658601726910950510/07658601726910950510/14lQCOt5gHB6lk9995hsgB5cxgH8BUMU1?e=download&nonce=dj7n4jm3ie44c&user=07658601726910950510&hash=orscqeqh98gop1iet5s6gqlov65qn8o1"
disease <- read.csv("USChronicDiseaseIndicators.csv", na.strings = "")
```
Getting an overview of the data
```{r}
glimpse(disease)
```
This data set has 519,718 observations and 34 variables. 
```{r}
head(disease)
```
```{r}
tail(disease)
```

Table of summary Statistics
```{r}
summary(disease)
```
Visualization of missing value pattern. 
From the data set we can visualize the missing value pattern. Read areas of the following graph indicated missing values. 
```{r}
missmap(disease)
```

From the summary statistics and missing value plot we can see that the data set has 37% missing values and most of them are in the following variables
```{r}
names(-which(colMeans(is.na(disease))>0.1))
```

# Count number of missing values in each column
```{r}
lapply(disease, function(x){sum(is.na(x))})
```


# Tidying and Transformation

Drop columns that have more than 10% missing
```{r}
disease_no_miss <- disease[,-which(colMeans(is.na(disease))>0.1)]
```
Dimension after dropping missing values
```{r}
dim(disease_no_miss)
```
After dropping missing value columns there are only 18 variables left. 

Get a glimpse of the data.

```{r}
glimpse(disease_no_miss)
```



# Exploratory Data Analysis

Top 5 location with chronic disease
```{r}
disease_no_miss %>%
    count(LocationDesc)%>%
    arrange(desc(n)) %>%
    head()
```
Bottom 5 location with chronic disease
```{r}
disease_no_miss %>%
    count(LocationDesc)%>%
    arrange(-desc(n)) %>%
    head()
```


Top 5 Data Source with chronic disease
```{r}
disease_no_miss %>%
    count(DataSource)%>%
    arrange(desc(n)) %>%
    head()
```

Bottom 5 Data Source with chronic disease
```{r}
disease_no_miss %>%
    count(DataSource)%>%
    arrange(-desc(n)) %>%
    head()
```

Bar plot of Chronic Diseases
```{r}
barplot(table(disease_no_miss$Topic), main = "Distribution of Chronic Disease Topics")
```

From the bar plot we can see that most of the chronic diseases are the following top 5
```{r}
disease_no_miss %>%
    count(Topic)%>%
    arrange(desc(n)) %>%
    head()
```

Visualization of data value type
```{r}
barplot(table(disease_no_miss$DataValueType))
```

Top data value types are the following
```{r}
disease_no_miss %>%
    count(DataValueType)%>%
    arrange(desc(n)) %>%
    head()
```

Stratification Category
```{r}
barplot(table(disease_no_miss$StratificationCategory1))
```

Data Value Type Plot
```{r}
barplot(table(disease_no_miss$DataValueTypeID))
```

# Findings
Real world data sets are often messy and have lot's of missing values. From the exploratory analysis we've found that Diabetes, Chronic Obstructive Pulmonary Disease, Cardiovascular Di seas, Arthritis are the most occurring chronic diseases. 
From the overall analysis we've learned that the chronic diseases are a major concern in the USA though the data set is missing several necessary information.
Arizona,Florida,Iowa have most chronic disease while Guam,Virgin Islands have least disease.
